---
classoptions: 
  - sn-basic      
  - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: Substitution of red meat with legumes and risk of primary liver cancer in 126,744 UK Biobank participants a prospective cohort study

authors: 
  - prefix:
    firstname: Niels
    particle:
    lastname: Bock
    suffix:
    degrees:
    email: nb@ph-au.dk
    affiliation: 1
    corresponding: TRUE

  - firstname: Fie
    lastname: Langmann
    degrees:
    affiliation: 1
    
  - firstname: Christina
    particle: C.
    lastname: Dahm
    affiliation: 1

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: Department of Public Health
      orgname: Aarhus University
    address:
        city: Aarhus
        country: Denmark
  
keywords:
  - Food Substitutions
  - Liver cancer
  - Red meat
  - Legumes

# Sample for structured abstract
abstract: |
  **Purpose**: Lifestyle-associated primary liver cancer is on the rise worldwide. Preventative strategies are warranted. We aimed to estimate the effect of substituting unprocessed red meat, processed red meat and total red meat with legumes on primary liver cancer in a free-living population.

  **Methods**: We analyzed data from 126,744 UK Biobank participants who completed >= 2 24-hour diet recall questionnaires. Substitution of 15g/day of legumes with 15/day of unprocessed red meat, processed red meat and total red meat was statistically modeled using the leave-one-out model. Baseline characteristics were collected from the initial assessment visit. Information on liver cancer diagnoses was collected via external linkage to inpatient hospital episodes or central cancer registries.

  **Results**: During a median follow-up time of 11.3 years, 173 participants developed liver cancer. In the fully adjusted models, no effect of substituting 15/day of legumes with unprocessed red meat (HR: 0.97 (95% CI 0.91-1.03);), processed red meat (HR: 1.02 (95% CI 0.93-1.13)) or total red meat (HR: 0.98 (95% CI 0.93-1.04)) was observed. The results were robust to sensitivity analyses.

  **Conclusion**: Overall, no association between substituting red meat with legumes and liver cancer was observed. Further research with longer follow-up time is warranted to increase the number of liver cancer events.

bibliography: bibliography.bib

output:
  bookdown::pdf_book:
    base_format: rticles::springer_article
    number_sections: false
    latex_engine: xelatex
    
header-includes:
  - \usepackage{pdflscape}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.pos = "b",
                      out.extra = "")
library(here)
library(gt)
library(gtsummary)
library(tidyverse)
library(survival)
library(ggconsort)
library(scales)
library(ggdag)
targets::tar_config_set(store = here::here("_targets"))
data <- targets::tar_read(data_main)
data_flowchart <- targets::tar_read(data_flowchart)
data_hcc <- targets::tar_read(data_with_hcc)
data_icc <- targets::tar_read(data_with_icc)
data_alc <- targets::tar_read(data_with_alc)
data_misreporter <- targets::tar_read(data_with_misreporter)
data_3_ques_comp <- targets::tar_read(data_with_3_ques_comp)
data_liver_disease <- targets::tar_read(data_without_liver_disease)
data_any_cancer <- targets::tar_read(data_without_cancer)
data_death <- targets::tar_read(data_with_death)
source(here("R","2-table1.R"))
source(here("R","4-diet-table.R"))
source(here("R","3-1-main-analysis-table.R"))
source(here("R","7-table-legume-quintile.R"))
source(here("R","6-table-cancer-type.R"))
source(here("R", "sens-table.R"))
```

# Introduction {#sec1}

The main aim of this study was to estimate the effect of substituting
unprocessed red meat, processed red meat and total red meat with legumes
on primary liver cancer in a free-living population.

# Research Design and Methods {#sec2}

## Study population {#subsec1}

The UK Biobank, a population-based prospective cohort, was initiated in
2006. [@sudlow2015] During 2006-2010, more than 500,000 participants,
aged 40-69, were recruited and visited designated assessment centres
across the UK Participants provided information about age, sex,
sociodemographic factors (education, Townsend deprivation index, living
alone) and lifestyle factors (smoking, alcohol consumption, physical
activity) via touch screen questionnaires and computer-assisted
interviews. Anthropometric data (waist circumference) were collected via
physical measurements [@RN113].

## Dietary assessment {#subsec2}

A web-based 24-hour dietary recall was administered at the end of the
initial assessment visit for the last 70,000 recruited participants
[@RN115]. From February 2011 to April 2012, 320,000 participants who had
provided an e-mail address were invited on four separate occasions to
complete the 24-hour dietary recall, the Oxford WebQ, of which 210,947
participants completed at least one. The Oxford WebQ covered 206 food
items and 32 beverage items commonly consumed in the UK. Intakes were
reported in standard units of measurements, e.g., servings, cups,
slices, etc. with intake categories ranging from 0 to 3+ units
[@piernas2021]. The Oxford WebQ has been validated against
interviewer-based 24-hour dietary recalls and biomarkers [@Liu2011;
@Greenwood2019].

Researchers defined79 food groups and 14 beverage groups from the Oxford
WebQ using the UK National Diet and Nutrition Survey categories
[@piernas2021]. These food and beverage groups were used when defining
the food groups used in the substitution analyses (Supplementary Table
1). Legumes were defined as dietary pulses, baked beans, tofu-based
products, peas, hummus, soy drinks, and soy-based desserts and yogurt.
Red meat intake was defined as intake of beef, pork, lamb, or other
meat, including offal. Processed red meat intake was defined as
sausages, bacon (with and without fat), ham, or liver pate. Other food
groups included were animal-based foods, unhealthy plant-based foods,
healthy plant-based foods, and alcoholic beverages (Supplementary Table
1). Animal-based and healthy and unhealthy plant-based food foods were
grouped based on plant-based diet indices from previous studies
[@Thompson2023; @Heianza2021; @Satija2017; @Satija2016]. An overview of
included foods in each food group is displayed in Supplementary Table 1.

Due to the incapability of a single 24-hour dietary recall to properly
assess habitual dietary intake and variation in diet over time
[@thompson2013; @gurinovic2017], only participants who completed two or
more Oxford WebQs were eligible for inclusion in this study.

## Liver cancer assessment {#subsec3}

Liver cancer was defined according to ICD-10 diagnosis codes C22.0 for
Hepatocellular carcinoma (HCC) or C22.1 for Intrahepatic
cholangiocarcinoma (ICC) and ICD-9 diagnosis codes 1550 Malignant
neoplasm of liver, primary or 1551 Malignant neoplasm of intrahepatic
bile ducts. Incident and prevalent cases of liver cancer and
corresponding diagnosis dates were obtained via external linkage to
central cancer registries or hospital inpatient episodes [@RN112;
@RN114].

## Assessment of confounders {#subsec4}

Confounders were defined *a priori* from a literature review of the
background literature and illustrated using directed acyclic graphs
(Supplementary Figure 1). The following confounding variables were
selected: age at baseline (years, continuous), sex (male, female),
educational level (high: College or University degree, intermediate: A
levels/AS levels, O levels/GCSEs, or equivalent, low: none of the
previous mentioned), Townsend Deprivation Index (continuous), Living
alone (yes, no), waist circumference (cm, continuous), physical activity
(above/below the 2017 UK Physical activity guidelines of 150 minutes of
moderate activity per week or 75 minutes of vigorous activity, or
unknown), smoking (pack years as a proportion of lifespan exposed to
smoking, continous), and alcohol intake (g/day, continuous). All
confounders except age were selected from the initial assessment visit
before the start of follow-up.

## The substitution model {#subsec5}

The substitution analyses were conducted by replacing an equal mass of
meat with legumes. The size of the substitution was set to 15 g of
legumes for 15 g of meat to keep the substitution size below the mean
intake of any of the substituted food groups in the cohort. The
substitutions were modelled using the leave-one-out-approach in which
variables for every food group along with a variable for total food
intake are included, except the food group that are to be substituted
[@Ibsen2021]. To estimate substitution of 15 g of all red meats (red and
processed) with 15 g of legumes, the following model was defined:

```{=tex}
\begin{align}
\log(h(t;x)) &= \log(h_0(t)) + \beta_1 \text{Legumes (15g)} + \beta_2 \text{Total food intake (g)} \nonumber \\
&\quad + \beta_3 \text{Other food groups (g)} + \beta_4 \text{Covariates}
\end{align}
```
\noindent When substituting only red meat with legumes, processed red
meat was added to the model:

```{=tex}
\begin{align}
\log(h(t;x)) &= \log(h_0(t)) + \beta_1 \text{Legumes (15g)} + \beta_2 \text{Processed red meat (15g)} \nonumber \\
&\quad + \beta_3 \text{Total food intake (g)} + \beta_4 \text{Other food groups (g)} \nonumber \\
&\quad + \beta_5 \text{Covariates}
\end{align}
```
\noindent When substituting only processed red meat with legumes, red
meat was added to the model:

```{=tex}
\begin{align}
\log(h(t;x)) &= \log(h_0(t)) + \beta_1 \text{Legumes (15g)} + \beta_2 \text{Red meat (15g)} \nonumber \\
&\quad + \beta_3 \text{Total food intake (g)} + \beta_4 \text{Other food groups (g)} \nonumber \\
&\quad + \beta_5 \text{Covariates}
\end{align}
```
## Statistical analysis {#subsec6}

Multivariable-adjusted Cox proportional hazards regression models were
used to estimate hazard ratios (HR) with corresponding 95% confidence
intervals (CI) with age as the underlying timescale. Participants were
followed from the date of their last completed Oxford WebQ until the
occurrence of the event of interest or due to right censoring, whichever
came first. Participants were right censored in the event of death, loss
to follow-up, or administrative end of follow-up (October 31, 2022). Two
levels of adjustments were added to the substitution model. Model 1 was
minimally adjusted for age, total weight of food intake, and all other
food groups to fit the substitution model. Model 2 was further adjusted
for sex, educational level, Townsend Deprivation Index, living alone,
physical activity, smoking, alcohol intake, and waist circumference.

In secondary analyses, each cancer type was analysed separately to
evaluate if the pooling of HCC and ICC as one outcome in the main
analysis was justified. Furthermore, to estimate the effect of legume
intake regardsless of other dietary components, legume consumers
(divided into quartiles) were compared to non-consumers.

To evaluate the robustness of the main analyses, sensitivity analyses
were performed on subsamples of participants by excluding those with
high alcohol intake (more than 32 grams per day for men and 24 grams per
day for women), implausible energy intake (≤ 3200 or ≥ 16800 kJ/day for
men and ≤ 2000 or ≥ 14000 kJ/day for women), any liver disease before
baseline, any type of cancer before baseline, and fewer than 3 completed
Oxford WebQs. As neither the central cancer registries nor the hospital
inpatient registries were complete, liver cancer diagnoses retrieved
from death registries were included in a sensitivity analysis to test
for outcome misclassification bias. Lastly, one of our causal
assumptions was that anthropometry confounded the causal relationship
between replacing red meat with legumes and liver cancer; however,
strong arguments exist giving support to anthropometry being a mediator
between diet and health outcomes. Thus, to test for erroneously
conditioning on a potential mediator, waist circumference was removed in
a sensitivity analysis. Sensitivity analyses were modelled like the
fully adjusted models in the main analyses.

All analyses were conducted in R (version 4.1.1) with a significance
level of 5 %.

# Results {#sec3}

After excluding participants with liver cancer before baseline,
participants lost to follow-up before baseline, and participants with
errors in the diet data, 126,744 participants remained who had completed
two or more diet questionnaires (Figure \ref{fig:fig1}).

```{r fig1, out.width = "100%", fig.cap = "Flowchart of included participants. Missing diet data was merged with loss to follow-up before baseline due to n being less than 5. It should be noted that not all UK Biobank participants were invited to complete an Oxfords WebQ. Only the last 70,000 participants to visit an assessment center were asked to complete an Oxford WebQ at the end of their visit. Further Oxfords WebQs were sent to ~320,000 participants who provied an e-mail adress."}
source(here("R", "flowchart.R"))
flowchart
```

During a median follow-up time of 11.3 years, 173 participants developed
liver cancer. Participants who developed liver cancer were older at
baseline, had a higher waist circumference, were less physically active,
fewer had never smoked, and more were male, compared to all included
participants (Table 1).

Mean daily energy intake and food intake and daily intake of all
specified food groups in grams are presented in table 2.

No association was found for substituting 15 g/day of legumes with 15
g/day of total red meat, unprocessed red meat, or processed red meat and
risk of primary liver cancer in model 1 (Table 3: total red meat:
`r inline_text(m1t, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
unprocessed red meat:
`r inline_text(m1r, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
processed red meat:
`r inline_text(m1p, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`).
The estimated associations changed minimally but remained
non-significant with further adjustments (Table 3: total red meat:
`r inline_text(m2t, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
unprocessed red meat:
`r inline_text(m2r, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
processed red meat:
`r inline_text(m2p, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`).

In secondary analyses, when analyzing the substitution association for
HCC and ICC separately, the risk of HCC was positively associated with
substituting total, unprocessed, or processed red meat with legumes
(Supplementary Table 2, total red meat:
`r inline_text(m2t_hcc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
red meat:
`r inline_text(m2r_hcc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
processed red meat:
`r inline_text(m2p_hcc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`).
The association between substituting total or unprocessed red meat with
legumes and ICC indicated inverse associations (Supplementary Table 2:
total red meat:
`r inline_text(m2t_icc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
red meat:
`r inline_text(m2r_icc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`;
processed red meat:
`r inline_text(m2p_icc, variable = legume_daily_15, pattern = "HR: {estimate}, 95% CI: {conf.low}- {conf.high}")`).
The magnitude or direction of associations were not significantly
different across strata of liver cancer types.

In the adjusted non-substitution analysis, a mean intake of 6.3 grams of
legumes per day was associated with a reduced risk of liver cancer,
compared to no intake
(`r inline_text(m2t_leg, variable = legume_category, level = "Q1", pattern = "HR: {estimate}, 95% CI: {conf.low}-{conf.high}")`);
however, no associations were observed with further increase in legume
intake (supplementary table 3).

In sensitivity analyses, excluding participants based on high alcohol
intake, implausible energy intake, any liver disease or cancer before
baseline, or fewer than 3 completed Oxford WebQs did not alter the
estimates in any statistically significant way. This also applied for
including death registries as a source of liver cancer events and
excluding waist circumference from the fully adjusted analysis
(Supplementary table 4).

# Discussion {#sec4}

## Summary of results {#subsec7}

In this study of UK Biobank participants, we found no effect of
replacing 15g/day of red meat with legumes on risk of primary liver
cancer. The estimates did not change significantly in any of the
sensitivity analyses. When stratifying by liver cancer type, replacing
total red meat and unprocessed red meat with legumes showed an inverse
association with ICC, though the results were not statistically
significant.

## Discuss results {#subsec8}

e.g, why estimates for HCC and ICC go in opposite direction. sensitivity
analyses

Contrary to our hypothesis, replacing processed red meat with legumes
was associated with a non-significant increase in risk of primary liver
cancer, with a greater effect size compared to unprocessed red meat.
This pattern persisted across all sensitivity analyses. However, the
estimates for processed red meat were labeled with less confidence due
to the low median intake. These findings align with recent analyses from
the UK Biobank, which indicated that unprocessed red meat intake was
associated with a non-significant increase in liver cancer risk, with a
greater effect size than processed meat (both white and red meat)
[@Knuppel2020]. This supports the notion that processed meat may not be
associated with liver cancer risk in this population.

Stratifying on cancer type revealed that replacing unprocessed red meat
with legumes to some extent was associated with a decreased risk of ICC,
contrary to HCC (Supplementary table 4). However, the robustness of
these results were not tested in further analyses as they were
associated with greater uncertainty.

## Other studies: substitution analyses and foods related to disease {#subsec9}

Discuss studies on legumes/red meat and liver cancer/other diseases.
Other substitutions

The literature on food substitutions, particularly in relation to liver
cancer, is sparse. However, a recent meta-anlysis of observational
studies found a non-linear dose-response relationship between legume
intake and liver cancer risk, with a protective effect observed between
intakes of 8 g/day to 40 g/day [@liu2023a]. This somewhat contrasts with
our findings where any increase above 6.3 g/day of legumes was not
associated with a decreased risk of liver cancer, compared to no legume
intake. A recent meta-analysis of observational studies no association
between red or processed meat intake and HCC [@Di2023].

## Strength and limitations {#subsec10}

This study had some limitations. First, None of the registries used to
determine a diagnosis of liver cancer were complete or up-to-date at the
time of analysis [@RN112]. Data from external providers, such as the NHS
England, NHS Central Register or National Records of Scotland, were
estimated to be mostly complete by the UK Biobank at various dates,
ranging from 31 December 2016 for cancer data from Wales to 31 October
2022 for hospital inpatient data from England [@RN114]. This could
introduce misclassification bias, as individuals with liver cancer may
not be identified as events. However, the estimates were robust in a
sensitivity analysis that included death registries as an additional
source of liver cancer diagnoses to accommodate missing outcome events.

Second, the relatively low number of events limited our ability to
adjust for confounding factors. Too many adjustment levels per event can
compromise the validity of the multivariable Cox regression model,
potentially causing biased estimates. To ensure statistical validity, we
aimed for at least 10 events per variable in the main analysis by
limiting the number of adjustment levels, using fewer and broader food
groups, and fewer levels for categorical covariates. This approach was
guided by our *a priori* causal assumptions. Although this method helped
maintain statistical validity, it may have increased residual
confounding by diluting the importance of specific food groups.
Additionally, risk factors that we could not adjust for, such as
aflatoxin B1, a known liver carcinogen, may have contributed to residual
confounding.

Strengths of this study are the large sample size and the prospective
longitudinal design, which enhance generalizability and establish
temporality. Estimates were robust to exclusion of participants with
fewer than three completed Oxford WebQs, indicating that at least two
completed Oxford WebQs were sufficient to correct for measurement
errors.

Selection bias: Incomplete registries regarding cancer. Participants who
completed ≥ 2 diet questionnaires are healthier than the those who
didn't?

Lavt bortfald.

kun 5% deltagelse.

Information bias: Underreporting of unhealthy habits? Non-differentiated
misclassification. Participants with unhealthy habits are
underreporting? differentiated misclassification.

## Perspectives {#subsec11}

# Conclusion {#sec5}

# Acknowledgements {#sec6}

\newpage

# Tables {#sec7}

```{r table_one}
table_one
```

\newpage

\noindent

```{r table 2 diet table}
diet_table
```

\newpage

```{r table 3 main results table, echo=FALSE, tab.cap = "wefwioejf"}
table_main
```

\newpage

# Supplemental Material {#sec8}

```{r fig2, out.width = "100%", fig.cap = "Directed acyclic graph (DAG) visualizing the hypothesised causal relationship between replacing meat with legumes and liver cancer based on assumptions of biasing paths. Red nodes represent confounders. Square nodes represent the minimal sufficient adjustment set for estimating the effect of replacing meat with legumes on liver cancer. Shadowed arrows represent biasing paths."}
source(here("R", "dag-code.R"))
dag
```

\newpage

```{r sup table 1 food table new}
source(here("R", "food-group-table.R"))
food_group_table
```

\newpage

```{r sup table 2 stratified on cancer type}
table_cancer_type
```

\newpage

```{r sup table 3 legume quartiles, echo=FALSE}
table_legume
```

\newpage

```{r, results='asis', echo=FALSE}
cat('\\begin{landscape}')
```

```{r sup table 4 table sensitivity analyses, echo=FALSE}
table_sens
```

```{r, results='asis', echo=FALSE}
cat('\\end{landscape}')
```

\newpage
